---
title: "Chi=test diagnostic and Extension"
author: "David M. Gregory, M.A."
date: "`r Sys.Date()`"
output: pdf_document
---



```{r}
library(haven)
  library(dplyr)
  library(tidyr)
library(RVAideMemoire)
library(survey)

```

```{r}
d <- read_dta("C:/Users/walks/Documents/GitHub/Indictment-of-Political-Science-/R/CMPS 2020_replication data.dta")
```




```{r}

wtd_tab <- function(data, var, wt = "os_weight")
   {data %>% filter(!is.na(.data[[var]])) %>% 
   group_by(value = as_factor(.data[[var]])) %>%
   summarise(n = n(), wtd_n = sum(.data[[wt]], 
   na.rm = TRUE), .groups = "drop" ) %>%mutate(wtd_pct = 
   round(100 * wtd_n / sum(wtd_n), 1))}


```

```{r}
t1_vars <- c("S2_Racer5", "Q803", "S2_Race_Prime", "S2_Nativer2", "S2_Nativer3", "S2_Nativer4", "S2_Nativer5", "Q797r5", "S11")



```


```{r}

subs <- list("Screener" = d%>% 
filter(S2_Racer5 == 1), "Primary" = d%>%filter(S2_Race_Prime ==5),
"Census" =d%>% filter(Q803 == 3)) 

```

```{r}
sapply(subs, nrow)
```
```{r}
t2_vars <- c("S2_M", "S6", "Q14", 
"Q77r1", "Q21", "Q265", "Q57",
"Q266", "Q271", "Q819")
t2_vars <- c("S2_M", "S6", "Q14",
             "Q77r1", 
             "Q21", "Q265", 
             "Q57", "Q266", 
             "Q271", "Q819")


```


```{r}

lapply(subs, function(x) table(x$Q803))

```

```{r}
agree <- table(Screener = d$S2_Racer5 == 1, Census  = d$Q803 == 3, useNA = "ifany")

agree
```

```{r}



d_long <- d %>%
transmute(id      = row_number(), 
screener = as.integer(S2_Racer5 == 1),
primary  = as.integer(S2_Race_Prime == 5),
census = as.integer(Q803 == 3)
) %>% 
pivot_longer(-id, names_to = "method", values_to = "identified")

d_long <- d_long %>%
  mutate(method = factor(method, levels = c("screener", "primary", "census")))

cochran.qtest(identified ~ method | id, data = d_long)

```
# Survey-weighted membership test 
# Cluster on respondent id to account for within-subject correlation

```{r}

des_long <- svydesign(ids = ~id, weights = ~os_weight, 
data = d_long %>% left_join(d %>% transmute(id = row_number(),
os_weight), by = "id"))

fit <- svyglm(identified ~ method, design = des_long, family = quasibinomial())
survey::regTermTest(fit, ~method)





```

```{r}
m <- d %>% mutate(f_screen = as.integer(!is.na(S2_Racer5) & S2_Racer5 == 1),
f_prime  = as.integer(!is.na(S2_Race_Prime) & S2_Race_Prime == 5),
f_census = as.integer(!is.na(Q803)          & Q803 == 3))
```

```{r}
disc <- m %>%
filter(f_screen != f_census) %>%
mutate(group = if_else(f_screen == 1, "screener_only", "census_only"))

des <- svydesign(ids = ~1, weights = ~os_weight, data = disc)
```
# Q14: presidential vote. Third-party, other, and none are too thin to stand alone.
## Q77r1: discussed politics. Three "no" variants (unlikely, might, definitely will)
## all indicate the respondent did not discuss politics in the reference period.
## Q265: cultural event attendance. Collapse 7-point frequency to 3.
## Q266: tribal election participation. Category 3 (always, 2.6 in census_only) is thin.

```{r}

disc <- disc %>% mutate(
Q14c = case_when(Q14 == 1 ~ 1L,   
 Q14 == 2 ~ 2L,    
Q14 %in% c(3,4,5,6) ~ 3L,   
TRUE ~ NA_integer_ ),

Q77r1c = case_when(
Q77r1 == 1 ~ 1L,   
Q77r1 == 2 ~ 2L,   
Q77r1 %in% c(3,4,5) ~ 3L,   
TRUE~ NA_integer_),
Q265c = case_when(
Q265 %in% c(1,2)    ~ 1L,   
Q265 %in% c(3,4)    ~ 2L,   
 Q265 %in% c(5,6,7)  ~ 3L, TRUE ~ NA_integer_
  ),
Q266c = case_when(
Q266 == 0  ~ 0L,   
Q266 == 1  ~ 1L,   
Q266 %in% c(2,3) ~ 2L, TRUE ~ NA_integer_ ))
```

# Rebuild the design on the modified data
```{r}
des <- svydesign(ids = ~1, weights = ~os_weight, data = disc)
```
# Re-run with collapsed variables replacing the originals
```{r, results = "hide"} 
t2_collapsed <- c("S2_M", "S6", "Q14c", "Q77r1c", "Q21", "Q265c", "Q57", "Q266c", "Q271", "Q819")
res <- lapply(t2_collapsed, function(v) {
out <- svychisq(as.formula(paste0("~", v, " + group")), design = des)
data.frame(var = v, statistic = unname(out$statistic), p = out$p.value)})
res <- do.call(rbind, res)
res$p_bh <- p.adjust(res$p, method = "BH")
res



```
# Q819: collapse "very well" and "somewhat well" into "speaks well."
# Q265c: collapse to binary. "Monthly or more" is too thin even after the first collapse
## The substantive distinction is speakers vs non-speakers, not degree of fluency within speakers.
```{r}
disc <- disc %>% mutate(
  
Q819c = case_when(
Q819 %in% c(1,2)  ~ 1L,   
 Q819 == 3 ~ 2L,   
Q819 == 4 ~ 3L,   
    TRUE  ~ NA_integer_
  ),
  
Q265b = case_when(
Q265 %in% c(1,2)      ~ 1L,  
Q265 %in% c(3,4,5,6,7) ~ 2L, TRUE ~ NA_integer_ ))
```

```{r}

des <- svydesign(ids = ~1, weights = ~os_weight, data = disc)

t2_final <- c("S2_M", "S6", "Q14c", "Q77r1c", "Q21", "Q265b", "Q57", "Q266c", "Q271", "Q819c")

```
# Confirm no expected cells below 5
# Weighted cross-tabs for the final collapsed versions of Q819 and Q265

```{r}
for (v in c("Q819c", "Q265b")) 
print(svytable(as.formula(paste0("~", v, " + group")), design = des) %>%
prop.table(margin = 2) %>% round(3))

```


# Final Rao-Scott tests

```{r}
res_final <- lapply(t2_final, function(v) {
out <- svychisq(as.formula(paste0("~", v, " + group")), design = des)
data.frame(var = v, statistic = unname(out$statistic), p = out$p.value)})
res_final <- do.call(rbind, res_final)
res_final$p_bh <- p.adjust(res_final$p, method = "BH")
res_final



```



















