---
title: " Foxworth, Raymond, Ellenwood, Cheryl, and Evans, Laura. 2024. Replication Data for 'Surveying Native Americans: Early Lessons from the CMPS, PS: Political Science & Politics. Harvard Dataverse. DOI: 10.7910/DVN/DIZ4ZB"
author: "David M. Gregory, M.A."
date: "`r Sys.Date()`"
output: pdf_document 
---

# Load packages
Loads haven for reading Stata .dta files and handling labelled variables, dplyr for the filter/group/summarise pipeline, and tidyr, suppresses start up messages so they do not appear in the knitted output.
```{r}
suppressPackageStartupMessages({
  library(haven)
  library(dplyr)
  library(tidyr)
})
 
```

# Load data
 haven::read_dta reads Stata's binary .dta format into R,
 preserving value labels and variable labels as attributes.
```{r}
d <- read_dta("CMPS 2020_replication data.dta")

```
# Weighted tabulation function 
groups by the variable's labelled categories, and returns each category's unweighted count alongside its survey-weighted count and weighted percentage 
```{r}
wtd_tab <- function(data, var, wt = "os_weight") {
  data %>%
    filter(!is.na(.data[[var]])) %>%
    group_by(value = as_factor(.data[[var]])) %>%
    summarise(
      n = n(),
      wtd_n = sum(.data[[wt]], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(wtd_pct = round(100 * wtd_n / sum(wtd_n), 1))
}
 

```

# TABLE 1: Native American ID Across Question Types
Iterates over the ten identification-related variables and prints a weighted tabulation of each, reproducing the published Table 1 distribution of how respondents are classified as Native under the screener, Census, primary-race, ancestry, and tribal-affiliation items
```{r}
t1_vars <- c(
  "S2_Racer5", "Q803", "S2_Race_Prime",
  "S2_Nativer2", "S2_Nativer3", "S2_Nativer4", "S2_Nativer5",
  "Q797r5", "Q798r5", "S11"
)

for (v in t1_vars) {
  cat("\n---", v, "---\n")
  print(wtd_tab(d, v))
}


```


# TABLE Characteristics by Identification Method
 Constructs three subsamples of Native respondents defined by three different identification instruments (screener S2_Racer5 == 1, primary-race S2_Race_Prime == 5, Census Q803 == 3), prints their unweighted sizes (1,070 / 713 / 656), then prints weighted tabulations of ten political and cultural variables within each subsample to reproduce Table 2. 

```{r}
subs <- list(
  "Screener" = d %>% filter(S2_Racer5 == 1),
  "Primary"  = d %>% filter(S2_Race_Prime == 5),
  "Census"   = d %>% filter(Q803 == 3)
)

cat("\nSample sizes (unweighted):\n")
for (nm in names(subs)) {
  cat("  ", nm, ":", nrow(subs[[nm]]), "\n")
}

t2_vars <- c(
  "S2_M", "S6", "Q14", "Q77r1", "Q21",
  "Q265", "Q57", "Q266", "Q271", "Q819"
)

for (v in t2_vars) {
  cat("\n---", v, "---\n")
  for (nm in names(subs)) {
    cat(" [", nm, "]\n")
    print(wtd_tab(subs[[nm]], v))
  }
}
 
```

# Diagnostic Tests: 
Cross-tabulates the screener Native flag against the Census Native flag over the full sample 


```{r}
agree <- table(
  Screener = d$S2_Racer5 == 1,
  Census   = d$Q803 == 3,
  useNA    = "ifany"
)
agree

 
```
# Mcnear Chi Square Test 
Confirms neither identification variable has missing values, cross-tabulates the screener Native flag against the Census Native flag, and runs McNemar's test on that 2x2 to test whether the screener-Census disagreement is directionally asymmetric rather than symmetric noise

```{r}
stopifnot(sum(is.na(d$S2_Racer5)) == 0, sum(is.na(d$Q803)) == 0)
agree <- table(Screener = d$S2_Racer5 == 1,
               Census   = d$Q803 == 3,
               useNA    = "ifany")
agree
mcnemar.test(agree, correct = FALSE)
 
```




